---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/cogsci/EM3/Assignment I/assignment1")

library(pacman)

p_load(tidyverse, stringr, simr, lmerTest)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
```{r}
da <- read.csv("pca_autist_data.csv", sep = ",") %>% 
  na.omit()

```


- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}


  mluModel=lmer(CHI_MLU ~ VISIT*Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 
  
  summary(mluModel)
  power_diagnosis_VISIT = powerSim(mluModel,fixed("VISIT:Diagnosis"),nsim=200)
  
  power_diagnosis_VISIT

  
#Assuimng that the estimate of Variable1 is x as estiamted b y the model, and the sample size is constant at Y, the power of the main effect is X (CONF1, CONF2).
```
```{r}

  power_diagnosis= powerSim(mluModel,fixed("Diagnosis"),nsim=200)
  
  power_diagnosis
```

```{r}

mluModel=lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 
power_visit= powerSim(mluModel,fixed("VISIT"),nsim=200)
  
power_visit
```

### Exercise 2


- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
```{r}
mluModel <- lmer(CHI_MLU ~ VISIT * Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 


fixef(mluModel)["VISIT"] <- 0.1
fixef(mluModel)["DiagnosisTD"] <- -0.1
fixef(mluModel)["VISIT:DiagnosisTD"] <- 0.1
powerCurveV = powerCurve(mluModel, fixed("VISIT:Diagnosis"),along="SUBJ", nsim=50)

visit_diag <- powerCurveV
plot(visit_diag)
#not enough power - see simulation of particiapnts
```

```{r}
mluModel <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 


fixef(mluModel)["VISIT"] <- 0.1
fixef(mluModel)["DiagnosisTD"] <- -0.1

powerCurveV = powerCurve(mluModel, fixed("Diagnosis"),along="SUBJ", nsim=50)

diag <- powerCurveV
plot(diag)
#not enough power - see simulation of particiapnts
```

```{r}
mluModel <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 


fixef(mluModel)["VISIT"] <- 0.1
fixef(mluModel)["Diagnosis"] <- -0.1

powerCurveV = powerCurve(mluModel, fixed("VISIT"),along="SUBJ", nsim=50)

visit <- powerCurveV
plot(visit)
#not enough power - see simulation of particiapnts
```



```{r}
pca_mluModel <- lmer(CHI_MLU ~ TC3 + TC4 + TC1 + (1 + VISIT | SUBJ), data =da,REML=F) 
  

fixef(pca_mluModel)["TC3"] <- 0.1
fixef(pca_mluModel)["TC4"] <- 0.1
fixef(pca_mluModel)["TC1"] <- -0.1
power_tc3 = powerCurve(pca_mluModel, fixed("TC3"),along="SUBJ", nsim=15)
  

plot(power_tc3)
```
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect

- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}





createNewData <- function (participants,visits,model){
  require(MASS)
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(Visit=1:visits,Child.ID=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$Child.ID]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma=cov(ranef(model)$SUBJ))
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$Visit + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}


mluModel <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1 + VISIT | SUBJ), data =da,REML=F) 
fixef(mluModel)["VISIT"] <- 0.1
fixef(mluModel)["DiagnosisTD"] <- -0.1

sim_data <- createNewData(500, 6, mluModel)

sim_mluModel <- lmer(CHI_MLU ~ Visit + Diagnosis + (1 + Visit | Child.ID), sim_data,REML=F) 

powerCurveV = powerCurve(sim_mluModel, fixed("Diagnosis"),along="Child.ID", nsim=50)

sim_visit_diag <- powerCurveV 
sim_plot1 <- plot(sim_visit_diag)


```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.

```{r}
#Simulating 30 kids. CreateNewData will automatically divide subjects equally between diagnoses.
thirty_sim <- createNewData(30,6, mluModel)

thirty_sim_model <- lmer(CHI_MLU ~ Visit*Diagnosis + (1 + Visit | Child.ID), thirty_sim,REML=F) 

thirty_interaction = powerSim(thirty_sim_model,fixed("Visit:Diagnosis"),nsim=200)
thirty_interaction
```
After 200 simulations we observe a power of 0% for our interaction effect. 

```{r}

thirty_sim_model <- lmer(CHI_MLU ~ Visit + Diagnosis + (Visit | Child.ID), thirty_sim,REML=F)
thirty_diagnosis = powerSim(thirty_sim_model,fixed("Diagnosis"),nsim=200)
thirty_diagnosis
```
Our main of effect of Diagnosis obtains a power of 34.5%

```{r}
thirty_visit = powerSim(thirty_sim_model,fixed("Visit"),nsim=200)
thirty_visit
```
