---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/cogsci/EM3/Assignment I/assignment1")
setwd("~/cogsci/EM3/Assignment I/assignment1")

library(pacman)

p_load(tidyverse, stringr, simr, crqa,tseriesChaos, lmerTest)
p_load(tseriesChaos, nonlinearTseries,SparseM, crqa)

```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
- Extract "complex" descriptors: recurrence quantification analysis
```{r}
filenames <- list.files(path = "pitch", full.names = T)
da <- read.delim(filenames[1])

plot(da$f0)

```

Plotting pitch development over time


Standard descriptors
```{r}
da %>% 
  summarise(mean_pitch = mean(f0),
            sd_pitch = sd(f0),
            range = min(f0) - max(f0))

```



Less standard descriptors
 (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
```{r}
da %>% 
  summarise(median_pitch = median(f0),
            mean_abso_devi = mad(f0),
            coef_of_var = sd(f0) / mean(f0))

            

```

RCQ
Calculating optimal delaay
```{r}

 
mutual(da$f0,lag.max = 50)

tau = 10
```
Decision rule for choosing tau: Choose first value that is followed by an increase: I judge that tau = 10

```{r}
plot(false.nearest(da$f0,m=5,d=15,t=0))

d <- 4
```

based on the false nearest neighbors I choose to keep 4 dimensions.
A radius of 15 is tried for the crqa analysis.

```{r}

Results=crqa (da$f0, da$f0, delay=tau, embed=d, radius=15,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

RP=Results$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols)


```

Extracting RQA measures
```{r}
Results$RR
Results$L
Results$DET

```
RR: The recurrence rate, i.e. the percentage of points in the phase space falling within the radius, is 7.2
L: The average length of lines is 4.6
DET: 86% of the recurrent points form diagnonal lines






2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)


Median parameters over all subjects are found:

```{r}

files <- list.files(path = "pitch", full.names = T)



#parameter space. Max number of dimensions for embedding is set to 15
par = list(lgM =  15, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 100, normalize = 0,  
           rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  
           recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 


#Parameter optimization is done for all files
result <- lapply(files[1:10], function(x) {
  df <- read.delim(x)
  param <- try(optimizeParam(df$f0,df$f0,par, min.rec= 3.5, max.rec= 4.5))
  })



#error trials are removed from list. Here it's exploited that errors have a different class
result[sapply(result, function(x) !is.list(x))] <- NULL  

#1200 participants remains.

#List of paremters are turned into df and median is found. 
median_params <- lapply(result, data.frame) %>%
  bind_rows() %>% 
  summarise_all(funs(median))



median_params
```

Median params are
Radius = 14.321
embed dim = 5
delay = 14

```{r}
radius = median_params[[1]]
embed = median_params[[2]]
delay = median_params[[3]]

#importing participant data and extracting subject nr, diagnosis and trial


crqa_result <- ldply(files, function(x) {
  df <- read.delim(x)
  
  crqa <- try(crqa(df$f0, df$f0, delay=delay, embed=embed, radius=radius, normalize=0,rescale=0,mindiagline = 2,minvertline = 2))d
  
  crqa_df <- as.data.frame(NA)
  
  if(length(crqa) > 4){
    crqa_df <- crqa_df %>% 
  mutate(Subject = str_extract_all(x, regex("\\d{3}")), #adding subject col
        Diagnosis = ifelse(str_detect(x,"D1"),"Schizophrenia","Control"), #diagnosis col
        Study = str_extract_all(x, "Study[[:digit:]]"), #Study col
        Trial = str_extract_all(x, "T[[:digit:]]+"),
        file = x,
        mean = mean(df$f0, na.rm = T),
        stdDev = sd(df$f0, na.rm = T),
        range = max(df$f0) - min(df$f0),
        median = median(df$f0),
        MeanAbsoluteDeviation = mad(df$f0),
        coefficientOfVariation = (sd(df$f0) / mean(df$f0)),
        rqa_rr = crqa[[1]],
        rqa_DET = crqa[[2]],
        rqa_maxL = crqa[[4]],
        rqa_L = crqa[[5]],
        rqa_ENTR = crqa[[6]],
        rqa_LAM = crqa[[8]],
        rqa_TT = crqa[[9]])
      
  }else{
      crqa_df <- crqa_df %>% 
  mutate(Subject = str_extract_all(x, regex("\\d{3}")), #adding subject col
        Diagnosis = ifelse(str_detect(x,"D1"),"Schizophrenia","Control"), #diagnosis col
        Study = str_extract_all(x, "Study[[:digit:]]"), #Study col
        Trial = str_extract_all(x, "T[[:digit:]]+"),
        file = x,
        mean = mean(df$f0, na.rm = T),
        stdDev = sd(df$f0, na.rm = T),
        range = max(df$f0) - min(df$f0),
        median = median(df$f0),
        MeanAbsoluteDeviation = mad(df$f0),
        coefficientOfVariation = (sd(df$f0) / mean(df$f0)),
        rqa_rr=NA,
        rqa_DET = NA,
        rqa_maxL = NA,
        rqa_L = NA,
        rqa_ENTR = NA,
        rqa_LAM = NA,
        rqa_TT = NA)
  }
  
  }) 

crqa_df <- crqa_result %>% 
  mutate(Subject = unlist(Subject),
         Study = unlist(Study),
         Subject = as.factor(Subject),
         Diagnosis = as.factor(Diagnosis),
         Study = as.numeric(str_extract_all(Study, "[[:digit:]]")),
         Trial = as.numeric(str_extract_all(Trial, "[[:digit:]]+"))) 
  crqa_df <- crqa_df[2:19]



#write.csv(crqa_df, "crqa_df.csv", row.names = F)

```

3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?

Random effects: We let the effect of time (Trial) have individual baseline and development. Since we have a study desgin where each patient is matched by a control, we take into account that the accuracy of this match can vary. We do that by letter the effect of diagnosis vary. Lastly, we account for a possibly difference is baseline effects of the different studies.

Of fixed effects Study and Diagnosis is included as well as an interaction term.

As outcome variables models for recurrence rate and mean length of recurrence are made.

```{r}
  setwd("~/cogsci/EM3/Assignment I/assignment1")
  
  
  crqa_df <- read.csv("crqa_df.csv")

str(crqa_df)

crqa_df <- crqa_df %>% 
  mutate(Subject = as.factor(Subject),
         Study = as.integer(Study))

rr_mdl <- lmer(rqa_rr ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(rr_mdl)




```

We see a significant effect of diagnosis on recurrence rate. People with schizophrenia have higher reccurence rate  (beta = 2.4, sd = 1.114, t(83)=2.64, p = 0.01). 
```{r}
DET_mdl <- lmer(rqa_DET ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(DET_mdl)

```
We see no significant effct of DET

```{r}
maxL_mdl <- lmer(rqa_maxL ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(maxL_mdl)
```
```{r}
ENTR_mdl <- lmer(rqa_ENTR ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(ENTR_mdl)
```

```{r}
TT_mdl <- lmer(rqa_TT ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(TT_mdl)
```

- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 

```{r}

p_load(tidyverse, stringr, lmerTest, MuMIn, psych, corrgram, ModelMetrics, caret)

folds <- createFolds(crqa_df, 5)
error <- matrix(nrow = length(folds), ncol = 1)

fixed_coefs = matrix(nrow = length(folds), ncol = 2)

colnames(fixed_coefs) <- c("intercept", "predictor3")

for (i in seq_along(folds)) {
  train_folds = filter(crqa_df, !(Subject %in% folds[[i]]))
  predict_fold = filter(crqa_df, Subject %in% folds[[i]])
  
  train_model <-   lmer(rqa_rr ~ Diagnosis + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)
  
  test_error <- predict(train_model, predict_fold, allow.new.levels = T)
  
  error[i] <- ModelMetrics::rmse(predict_fold$CHI_MLU,test_error)
  
  fixed_coefs[i,] <- fixef(train_model) 
  
}
fixed_coefs
mean(error, na.rm = T)



```


3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?

```{r}

crqa_df <- crqa_df %>% 
  mutate(Subject = as.factor(Subject))

rr_mdl <- lmer(rqa_rr ~ Study + (Diagnosis + Trial| Subject)+ (1|Study), crqa_df)

summary(rr_mdl)
```



```{r}
rr_mdl <- lm(rqa_rr ~ Study , crqa_df)

summary(rr_mdl)

```

```{r}
rr_mdl <- lmer(scale(rqa_rr) ~ Study*Diagnosis + (Diagnosis + Trial| Subject) + (1|Study), crqa_df)

summary(rr_mdl)
```


4. Bonus Question: Compare effect size of diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time